Introducing the \NAME framework\\ for efficient Intrusion Detection\\ in the Internet of Things

### 2014-10-06 paper.tex
roughly updated title/abstract/title after meeting

@@

\title{Efficiently Detecting Intrusions\\in the Internet of Things}

@@

Supporting multiple intrusion detection (ID) algorithms imposes significant
overhead on Internet of Things (IoT) devices. Fusion of the underlying
algorithms can alleviate this, yet, proves to be time-consuming, repetitive and
error-prone. To address this problem we introduce id-foo, a framework for the
development of efficient intrusion detection systems (IDS). It consists of a
domain specific language (DSL) that allows formally describing ID algorithms on
a functional level. The id-foo code generator then fuses these algorithms and
produces more optimally organised source code. Integration of applications with
this code further enables reacting on threats as they occur, not post-mortem. A
side-by-side comparison shows that generated code reduces memory footprint,
message passing overhead and execution time in comparison to a straightforward
sequential implementation of the same algorithms.

@@

Along with all the wonderful possibilities, the IoT clearly holds an enormous
threat: by opening our homes and lives to all these interconnected devices, we
also open them to everyone who is able to break these new, virtual locks, often
without leaving a trace.

@@

% manet: local ids + opportunity of integration

In mobile ad hoc networks (MANET), which make up an important part of the IoT,
it is not possible to have this single point in the network that can oversee
all traffic \cite{mishra2004intrusion}. Every node is individually
approachable. If this outer security perimeter diminishes, every device has to
implement its own lines of defence. The services running on those devices can
take advantage of these local security layers. If the IDS detects a problem,
the service can immediately terminate the communication with the attacker,
before it's actually abused.

% manet: limited resources

Resources are another important factor: IoT devices typically have limited
batteries, little processing power and memory is restricted to the bare minimum.

@@

% my approach merged into scientific contribution

@@

The first contribution of this paper is a framework that enables the formal
description of ID algorithms on a functional level. An automated code
generation process produces corresponding source code, for a given platform and
configuration.

@@

% 3. integration

The final contribution is the integration of the generated IDS with the actual
software. Information from the different algorithms is unified and made
available to the applications, that can use it to actively protect themselves.

@@

\section{Evaluation}
\label{evaluation}

@@

\section{Implementation and Evaluation}

>>>>

% until we settle on the final name ;-)
\usepackage{xspace}
\newcommand{\NAME}{id-foo\xspace}

@@

\title{
Introducing the \NAME framework\\
for efficient Intrusion Detection\\
in the Internet of Things
}

@@

Supporting multiple intrusion detection (ID) algorithms imposes significant
overhead on Internet of Things (IoT) devices. Fusion of the underlying
algorithms can alleviate this, yet, proves to be time-consuming, repetitive and
error-prone. To address this problem we introduce \NAME, a framework for the
development of efficient intrusion detection systems (IDS). It consists of a
domain specific language (DSL) that allows formally describing ID algorithms on
a functional level. A code generator then fuses these algorithms and produces
more optimally organised source code. This way, \NAME addresses the
heterogeneity of both the IoT and ID, as well as the limited resources of IoT
devices. A side-by-side comparison shows that generated code reduces memory
footprint, message passing overhead and execution time in comparison to a
straightforward sequential implementation of the same algorithms.

@@

Along with all the wonderful possibilities, the IoT clearly holds a significant
threat: by opening our homes and lives to all these interconnected devices, we
also open them to everyone who is able to break these new, virtual locks, often
without leaving a trace.

@@

% manet: local ids -> localize -> limited resources

In mobile ad hoc networks (MANET), which make up an important part of the IoT,
it is not possible to have this single point in the network that can oversee
all traffic \cite{mishra2004intrusion}. If the outer security perimeter
diminishes, every device has to implement its own lines of defence.
Localisation of the IDS requires resources, and resources are another important
factor to consider: IoT devices typically have limited batteries, little
processing power and memory is restricted to the bare minimum.

@@

% my approach merged into ...
% scientific contribution

@@

The first contribution of this paper is a framework that enables the formal
description of ID algorithms on a functional level. An automated code
generation process produces corresponding source code, for a given platform and
configuration. The way, the framework addresses the heterogeneous nature of
both the IoT as a hardware and software platform, and that of the different ID
algorithms.

@@

@@

\section{Implementation and Evaluation}
\label{evaluation}

@@

Integration of applications with generated IDS allows actively and dynamically
responding to changes in the reputation of other nodes.

The services running on those devices can take advantage of these local
security layers. If the IDS detects a problem, the service can immediately
terminate the communication with the attacker, before it's actually abused.
